Community-Conditioned Language Models

### 2020-11-23 paper/paper.tex
finish float refactor

\begin{table*}
  \small
  \centering
  \input{floats/community_results.tex}
  \caption{
    Model performance by community. We show the absolute perplexity 
    of the baseline models and the information gain of the CCLMs.
    Note that the rows are sorted by the perplexity of the baseline
    LSTM model.
}
  \label{tab:community-results}
\end{table*}

@@

\begin{figure*}
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1}
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

@@

\begin{figure}
  \pairwisecommsim{lstm-3-1} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different categories (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

Indiscernibility of a given community is a robust measure: it is
almost perfectly correlated across CCLMs (such as transformers, LSTM
and various values of $l_c$).  In fact, we observe a correlation
between the indiscernibility measures by every pair of models with
Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
\cref{fig:lmcc_ppl}).

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
    \end{axis}
  \end{tikzpicture}
  \caption{%
    Correlation between indiscernibility for 
    the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc_ppl}
\end{figure}

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

@@

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Mean community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm_lmcc_ppl}
\end{figure*}

@@

On the other hand, community indiscernibility is correlated with
the information gain of the CCLM 
on messages from that community's test set 
(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm_lmcc_ppl}).
Communities for which the language model benefits most from 
the conditioned model are generally easier for the LMCC to classify.

@@

We cannot exclude a correlation (\cref{tab:cclm_lmcc_ppl}) between the perplexity of the
% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46)\jp{make consistent with data once the plot is stablilized}, so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.

>>>>

\begin{table*}
  \centering
  \input{floats/community_results.tex}
  \caption{
    Model performance by community. We show the absolute perplexity 
    of the baseline models and the information gain of the CCLMs.
    Note that the rows are sorted by the perplexity of the baseline
    LSTM model.
}
  \label{tab:community-results}
\end{table*}

@@

\begin{figure*}
\begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    legend style={anchor=north, legend columns=1, font=\sffamily\tiny},
  ]
    \nextgroupplot
    \commscatter{lstm-3-1_pca1}{lstm-3-1_pca2}
    \nextgroupplot
    \commscatter{transformer-3-3_pca1}{transformer-3-3_pca2}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
\end{tikzpicture}
\caption{Community embedding PCA of the best LSTM (left, $c=1$) and transformer (right, $c=3$) models.}
\label{fig:comm-pca}
\end{figure*}

@@

\begin{figure}
  \pairwisecommsim{snap_cos_sim}{lstm-3-1_cos_sim} 
  \caption{Pairwise cosine similarity between community vectors for 
    the membership-based embedding (x-axis) and
  the LSTM CCLM embedding with $c=1$ (y-axis). 
  Pairs from the same category are colored according to \cref{fig:comm-pca}.
  We also mark pairs from different categories (\texttimes) and pairs where one or both
  communities belong to the ``Other'' category (\textbigcircle).
  }
  \label{fig:pairwise-comm-sim}
\end{figure}

@@

Indiscernibility of a given community is a robust measure: it is
almost perfectly correlated across CCLMs (such as transformers, LSTM
and various values of $l_c$).  In fact, we observe a correlation
between the indiscernibility measures by every pair of models with
Pearson's $r \geq 0.98$ and $p < 0.0001$ (see, for example,
\cref{fig:lmcc-ppl}).

\begin{figure}
  \begin{tikzpicture}
    \begin{axis}
    \commscatter{lstm-3-1_indisc}{transformer-3-3_indisc}
    \end{axis}
  \end{tikzpicture}
  \caption{%
    Correlation between indiscernibility for 
    the best LSTM ($c = 1$) and transformer ($c = 3$) models
    (Pearson's $r = 0.99$, $p < 0.0001$).
  }
  \label{fig:lmcc-ppl}
\end{figure}

However, we observe that community-wise perplexity is only weakly (if at all) 
correlated with community indiscernibility under the corresponding latent 
LMCC (see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
In other words, the linguistic complexity of a given community
is a poor predictor of how distinct its language is.

@@

\begin{figure*}
  \begin{tikzpicture}
  \begin{groupplot}[
    group style={group size=2 by 1},
    yticklabels={,,}, xticklabels={,,},
    legend style={anchor=north, legend columns=1, font=\tiny},
    every node near coord/.append style={font=\sffamily\tiny}%
  ]
    \nextgroupplot
    \commscatter[]{lstm-3-1_lm_ppl}{lstm-3-1_indisc}
    \nextgroupplot
    \commscatter[]{lstm-3-1_info_gain}{lstm-3-1_indisc}
    \legend{General interest,Videogames,Technology,Sports,Female-focused,Other}
  \end{groupplot}
  \end{tikzpicture}
  \caption{%
    Mean community indiscernibility (y-axis), 
    in relationship to linguistic complexity (left) and 
    CCLM information gain (right). Shown here for the LSTM
    model with $l_c=1$. See \cref{tab:ind-corrs} for 
    Pearson's r correlations for all models.
  }
  \label{fig:cclm-lmcc-ppl}
\end{figure*}

@@

On the other hand, community indiscernibility is correlated with
the information gain of the CCLM 
on messages from that community's test set 
(see \cref{tab:ind-corrs} and, for example, \cref{fig:cclm-lmcc-ppl}).
Communities for which the language model benefits most from 
the conditioned model are generally easier for the LMCC to classify.

@@

We cannot exclude a correlation (\cref{tab:ind-corrs}) between the perplexity of the
% \bn{I'm not sure if we can make this claim or not with the Ppl-Ppl correlations. However, given that *all* of them have positive correlation, even though the p-values are somewhat higher than .1 maybe we can?}
CCLM and the indiscernibility of the corresponding community
\ref{hyp:rich-harder-to-identify}. Because we have relatively few,
and only large communities in our test set, it is difficult to
provide a conclusive answer, but this kind of correlation would suggest that
community-specific complexity can be partially attributed to using
more varied patterns. However, even the most complex communities
(perplexity per word above 100) have an indiscernibility of about 3
(out of a maximum of 46)\jp{make consistent with data once the plot is stablilized}, so they retain fairly specific patterns:
they are not simply an even mixture of the kind of comments found in
other communities.

### 2020-11-23 paper/plot_macros.tex
finish float refactor

@@

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    x tick label style={font=\sffamily\tiny,rotate=45,anchor=east},
    xtick=data,
    legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

\newcommand{\pairwisecommsim}[1]{%
  \begin{tikzpicture}
    \begin{axis}[]
      \addplot[
        scatter, only marks,
        font=\tiny,
        scatter src=explicit symbolic,
        scatter/classes={
          general-interest={mark=*, blue},
          videogames={mark=*, red},
          technology={mark=*, green},
          sports={mark=*, cyan},
          female-focused={mark=*, magenta},
          other={mark=o, black},
          different={mark=x}
        },
        ] table [x=snap, y=#1, meta=meta] {floats/comm_sim.csv};
    \end{axis}
  \end{tikzpicture}
}

>>>>

\usepackage{lmodern}

@@

\newcommand{\confusiongroupplot}[2]{%
  \begin{groupplot}[
    group style={group size=2 by 1}, 
    enlargelimits=false,
    axis line style={draw=none},
    %axis equal,
    xmin=-1, xmax=46,
    ymin=-1, ymax=46,
    width=8cm,
    xticklabels={Advice,AskWomen,BabyBumps,CFB,Drugs,EDH,EarthPorn,Fantasy,GameDeals,GlobalOffensive,Jokes,Kappa,KerbalSpaceProgram,KotakuInAction,LifeProTips,MLS,MMA,MaddenUltimateTeam,TwoXChromosomes,Warframe,airsoft,bodybuilding,breakingmom,cars,cringe,eu4,exjw,explainlikeimfive,femalefashionadvice,food,heroesofthestorm,jailbreak,justneckbeardthings,oculus,pcmasterrace,photography,reddevils,relationships,rupaulsdragrace,stopdrinking,streetwear,techsupport,todayilearned,toronto,videos,xxfitness},
    yticklabels=none,
    x tick label style={font=\sffamily\tiny,rotate=90,anchor=east},
    xtick=data,
    legend style={font=\sffamily\small},
    ]
    \nextgroupplot
    \confusionplot{#1}
    \nextgroupplot[colorbar right]
    \confusionplot{#2}
  \end{groupplot}
}

\newcommand{\pairwisecommsim}[2]{%
  \begin{tikzpicture}
    \begin{axis}[]
      \addplot[
        scatter, only marks,
        font=\tiny,
        scatter src=explicit symbolic,
        scatter/classes={
          general-interest={mark=*, blue},
          videogames={mark=*, red},
          technology={mark=*, green},
          sports={mark=*, cyan},
          female-focused={mark=*, magenta},
          other={mark=o, black},
          different={mark=x}
        },
        ] table [x=#1, y=#2, meta=category] {floats/comm_comm.csv};
    \end{axis}
  \end{tikzpicture}
}
